\texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles

### 2023-05-22 Manuscript/manuscript.tex
Add LCT's comment + funding information.

Adding a third dimension to 2D plots has been a long-standing challenge in visualization \cite{odonoghue_2010}. Our proposal addresses this fundamental challenge by introducing simple but effective design principles. Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economic solution. In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data. Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020}, to avoid possible overplotting, the proposed visualization can be also applied to visualized image-based spatially resolved data \cite{caicedo_2017}, in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).

@@

\subsection*{Funding}
This project was supported by the National Institute of Mental Health R01MH126393 (BG, SCH), and CZF2019-002443 (SCH) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation. All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.

>>>>

Adding a third dimension to 2D plots has been a long-standing challenge in visualization \cite{odonoghue_2010}. Our proposal addresses this fundamental challenge by introducing simple but effective design principles. These principles lead to visually easier-to-interpret graphics compared to geometry annotations (\textbf{Figure~\ref{fig:embedding}}). Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economic solution. In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data. Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020}, to avoid possible overplotting, the proposed visualization can be also applied to visualized image-based spatially resolved data \cite{caicedo_2017}, in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).  \texttt{escheR} could also be applied with kernel gene-weighted density plots from \texttt{Nebulosa} \cite{alquicirahernandez_2021} and other ggplot2-based visualizations.

@@

\subsection*{Funding}
This project was supported by the National Institute of Mental Health [R01MH126393 to BG and SCH, U01MH122849 to LAHM and LCT]; the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation[CZF2019-002443 to SCH]; the Lieber Institute for Brain Development to LAHM and LCT; and Aligning Science Across Parkinson’s [ASAP-000478, ASAP-000509 to MGP] through the Michael J. Fox Foundation for Parkinson’s Research.  All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.
